---
title: "Airline Passenger Satisfaction (Maven Analytics)"
author: "Paul Williams"
date: "`r Sys.Date()`"
output: html_document
---

# Airline Passenger Satisfaction

- https://yihui.org/knitr/options/#chunk_options
- https://kbroman.org/knitr_knutshell/pages/Rmarkdown.html

## Intro

The dataset and questions were sourced from Maven Analytics:  https://www.mavenanalytics.io/data-playground?page=2&pageSize=5

- Which percentage of airline passengers are satisfied?
- Does satisfaction vary by customer type?
- Does satisfaction vary by type of travel?
- What is the customer profile for a repeating airline passenger?
- Does flight distance affect customer preferences or flight patterns?
- Which factors contribute to customer satisfaction the most? What about dissatisfaction?

## Setup

```{r setup-chunks, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  warning=FALSE,
  message=FALSE,
  dev="svg"
)
```

## Installing packages

```{r install-packages, eval=FALSE, include=FALSE}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("svglite") # apt: libfontconfig1-dev
```

## Load data

```{r load-data}
library("readr")

data <- read_csv("../data/airline_passenger_satisfaction.csv")

head(data)
```

## Preview

```{r preview}
data %>%
  select(Satisfaction) %>%
  group_by(Satisfaction) %>%
  summarise(
    satisfaction_category = n()
  ) %>%
  print()
```

## Which percentage of airline passengers are satisfied?



```{r general-satisfaction-preview, include=FALSE}

total_survey_responses <- nrow(data)

calc_satisfaction_percentage <- function(satisfaction_totals) {
  one_response <- 100.0 / total_survey_responses
  group_total <- one_response * satisfaction_totals
  percentage <- round(group_total, 2)
  return (percentage)
}

satisfaction_summary <- data %>%
  select(Satisfaction) %>%
  group_by(Satisfaction) %>%
  summarise(satisfaction_totals = n()) %>%
  mutate(percentage = calc_satisfaction_percentage(satisfaction_totals))

print(satisfaction_summary)
```

```{r general-satisfaction-pie-chart}
library('dplyr')
library('ggplot2')
library('svglite')

pie_chart <- function(data, percent, name, title) {
  y=enquo(percent)
  fill=enquo(name)
  
  plot <- data %>%
    ggplot(aes(x="", y=!!y, fill=!!fill)) +
    geom_col(color="black") +
    geom_text(
      aes(label=paste(!!fill, "\n", !!y, "%")),
      position=position_stack(vjust=0.5)
    ) +
    coord_polar(theta="y") +
    theme_void() +
    theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title=title)
  
  return (plot)
}

satisfaction_summary %>%
  pie_chart(
    percent=percentage,
    name=Satisfaction,
    title="Majority of passengers are either neutral or disatisfied with their flight"
  ) +
  scale_fill_manual(values = c(
    "#ef7c8e",
    "#b6e2d3"
  ))

if (!dir.exists(file.path("..", "plots"))) {
  dir.create(file.path("..", "plots"))
}

ggsave("../plots/plot.svg")
```
